---
title: "DA5030.P2.Parpattedar"
author: "Shruti Parpattedar"
date: "March 17, 2019"
output: pdf_document
---
Note : Collaborated with Shantanu Bafna on Problem 2
# Problem 1

## Question 1
Import the data and add headers to the data frame.
```{r Question1.1}
data <- read.csv("income.csv", header = FALSE, fileEncoding = "UTF-8-BOM")
names(data) <- c("Age", "WorkClass", "fnlwgt", "Education", "Edu-num", "Marital-status", "Occupation", "Relationship", "Race", "Sex", "Cap-gain", "Cap-loss", "HrsPerWeek", "NativeCountry", "Income")
str(data)
```

## Question 2
After exploring the data, I noticed that most of the variables are almost normally distributed. 

ANother observation was that fnlwgt is right skewed, so we can apply the log transform to rectify the skew.
```{r Question1.2}
library(psych)
pairs.panels(data[,1:7])
pairs.panels(data[,8:15])
hist(data$fnlwgt)
hist(log(data$fnlwgt))
```

## Question 3
Using the categorical variables to create the frequency and likelihood tables.

Next, implementing a Naive Bayes Classifer
```{r Question1.3}
library(gmodels)
library(dplyr)

# Column numbers for the categorical variables
cat_variables <- c(2,4,6,7,8,9,10,14,15)

# Dataset with only categorical variables
cat_data <- data[, cat_variables]

createLikelihoodT <- function(dataset)
{
  n <- ncol(dataset) - 1 # excluding the Income column from the count
  
  # Creating the initial frequency and likelihood tables using only the first categorical variable 
  # with the income variable
  # firstTable <- CrossTable(dataset$Income, dataset[, 1])
  freqTable <- table(dataset$Income, dataset[, 1])
  # freqTable <- firstTable$t
  likelihoodTable <- matrix(c(freqTable[1,]/sum(freqTable[1,]), freqTable[2,]/sum(freqTable[2,])), nrow = 2, 
                            byrow = TRUE)
  colnames(likelihoodTable) <- colnames(freqTable)
  
  # Cbinding the frequency and likelihood tables for the remaining categorical variables to the intial tables 
  # for(i in 2:n)
  # {
  #   newTable <- CrossTable(dataset$Income, dataset[,i])
  #   freqTable <- cbind(freqTable, newTable$t)
  #   likelihoodTable <- cbind(likelihoodTable, newTable$prop.row)
  # }
  # return(likelihoodTable)
  
  for(i in 2:n)
  {
    new_freq_Table <- table(dataset$Income, dataset[,i])
    freqTable <- cbind(freqTable, new_freq_Table)
    new_LH_Table <- matrix(c(new_freq_Table[1,]/sum(new_freq_Table[1,]), new_freq_Table[2,]/sum(new_freq_Table[2,])), nrow = 2, byrow = TRUE)
    colnames(new_LH_Table) <- colnames(new_freq_Table)
    likelihoodTable <- cbind(likelihoodTable, new_LH_Table)
  }
  return(likelihoodTable)
}

likelihoodTable_orig <- createLikelihoodT(cat_data)

# Implementing the Naive Bayes classifier that returns the probability of the income being less than 50k
naiveBayes <- function(values, LH_Table)
{
  # P(<=50k)*product of P(<=50k and parameters)
  P_le50k <- table(data$Income)[1]*prod(LH_Table[1, values])
  
  # P(>50k)*product of P(>50k and parameters)
  P_gt50k <- table(data$Income)[2]*prod(LH_Table[2, values])
  
  result <- P_le50k / (P_le50k + P_gt50k)
  
  # Returning probability of income being <=50k using the values calculated above
  return(ifelse(result < 0.5, ">50k", "<=50k"))
}

```

## Question 4
Using the classifier to predict the outcome for the new data
```{r Question1.4}
vals <- c("White", "Female", "Federal-gov", "Bachelors", "India")
output <- naiveBayes(vals, likelihoodTable_orig)
paste("Result is", output)
```

## Question 5
10-fold cross validation on the classifier
```{r Question1.5}
library(caret)
folds <- createFolds(data$Income)

nb_results <- lapply(folds, function(x)
{
  train_data <- cat_data[-x,]
  test_data <- cat_data[x,]
  likelihood_train <- createLikelihoodT(train_data)
  test_res <- data.frame(Prediction = rep(" ", nrow(test_data)), stringsAsFactors = FALSE)
  
  for(i in 1:nrow(test_data))
  {
    test_res$Prediction[i] <- naiveBayes(unlist(test_data[i,]), likelihood_train)[[1]]
  }
  
  # Confusion matrix
  conf_mat <- table(test_data$Income, test_res$Prediction)
  
  # assigning the accuracy of this model to the vector
  sum(diag(conf_mat))/sum(conf_mat)

})

paste("Accuracy =", mean(unlist(nb_results)))

```

# Problem 2

## Question 1
Finding and eliminating outliers
```{r Question2.1}
data_Q2 <- read.csv("uffidata.csv", stringsAsFactors = FALSE)
data_Q2 <- data_Q2[-100,]
data_Q2$Sale.Price <- as.numeric(gsub(",", "",data_Q2$Sale.Price))
str(data_Q2)
bsmnt_mean   <- mean(data_Q2$Bsmnt.Fin_SF)
bsmnt_sd     <- sd(data_Q2$Bsmnt.Fin_SF)
bsmnt_zscore <- (bsmnt_mean - data_Q2$Bsmnt.Fin_SF) / bsmnt_sd

o <- which(bsmnt_zscore > 1)
data2 <- data_Q2[-o,]

Lot.Area_mean   <- mean(data2$Lot.Area)
Lot.Area_sd     <- sd(data2$Lot.Area)
Lot.Area_zscore <- (Lot.Area_mean - data2$Lot.Area) / Lot.Area_sd
o <- which(Lot.Area_zscore > 1)
data3 <- data2[-o,]

final_data <- data3
```

## Question 2
Full correlation matrix
```{R Question2.2}
cor(final_data)
```

## Question 3
Multiple regression model for Sales Price
```{r Question2.3}
model_Q2 <- lm(Sale.Price ~ ., final_data)
summary(model_Q2)

model2_Q2 <- lm(Sale.Price ~ Year.Sold + Enc.Pk.Spaces + Living.Area_SF + Pool, final_data)
summary(model2_Q2)
```

## Question 4
UFFI and change in the value of a property
```{r}
UFFI.glm <- glm(formula = UFFI.IN ~ Sale.Price, 
                data = final_data, 
                family = binomial)

newdata <- data.frame(Sale.Price = 128304.2)

predict(UFFI.glm, newdata, type = "response")
```

## Question 5
Predicting Sales price for the given data along with its 95% CI.
```{r Quetion2.5}
model1 <- lm(Sale.Price ~ X45.Yrs. + Bsmnt.Fin_SF + Brick.Ext + Lot.Area + Enc.Pk.Spaces + Living.Area_SF + UFFI.IN + Central.Air + Pool, data = data_Q2)

sm1 <- summary(model1)

X45      <- 1
Basement <- 0
Brick    <- 0
LotArea  <- 7800
Parking  <- 1
Living   <- 1720
UFFI     <- 1
Central  <- 0
Pool     <- 0
salePrice <- sm1$coefficients[[1]] + sm1$coefficients[[2]]*X45 + sm1$coefficients[[3]]*Basement + sm1$coefficients[[4]]*Brick + sm1$coefficients[[5]]*LotArea + sm1$coefficients[[6]]*Parking + sm1$coefficients[[7]]*Living + sm1$coefficients[[8]]*UFFI + sm1$coefficients[[9]]*Central + sm1$coefficients[[10]]*Pool

paste("Predicted Value =", salePrice)
paste("95% CI =", salePrice - 1.96*22230, salePrice + 1.96*22230)
```


# Problem 3

## Question 1
Importing data
```{r Question3.1}
data_Q3 <- read.csv("bank-full.csv", sep = ";")

validation_Q3 <- read.csv("bank.csv", sep = ";")
```

## Question 2
Creating dummy codes.

Next, implementing a logistic regression model.
```{r Question3.2}
data_Q3$job1 <- ifelse(data_Q3$job == "admin.", 1, 0)
data_Q3$job2 <- ifelse(data_Q3$job == "unknown", 1, 0)
data_Q3$job3 <- ifelse(data_Q3$job == "unemployed", 1, 0)
data_Q3$job4 <- ifelse(data_Q3$job == "management", 1, 0)
data_Q3$job5 <- ifelse(data_Q3$job == "housemaid", 1, 0)
data_Q3$job6 <- ifelse(data_Q3$job == "entrepreneur", 1, 0)
data_Q3$job7 <- ifelse(data_Q3$job == "student", 1, 0)
data_Q3$job8 <- ifelse(data_Q3$job == "blue-collar", 1, 0)
data_Q3$job9 <- ifelse(data_Q3$job == "self-employed", 1, 0)
data_Q3$job10 <- ifelse(data_Q3$job == "retired", 1, 0)
data_Q3$job11 <- ifelse(data_Q3$job == "technician", 1, 0)

data_Q3$marital1 <- ifelse(data_Q3$marital == "married", 1, 0)
data_Q3$marital2 <- ifelse(data_Q3$marital == "divorced", 1, 0)

data_Q3$edu1 <- ifelse(data_Q3$education == "secondary", 1, 0)
data_Q3$edu2 <- ifelse(data_Q3$education == "primary", 1, 0)
data_Q3$edu3 <- ifelse(data_Q3$education == "tertiary", 1, 0)

data_Q3$default10 <- ifelse(data_Q3$default == "yes", 1, 0)

data_Q3$housing10 <- ifelse(data_Q3$housing == "yes", 1, 0)

data_Q3$loan10 <- ifelse(data_Q3$loan == "yes", 1, 0)

data_Q3$contact1 <- ifelse(data_Q3$contact == "telephone", 1, 0)
data_Q3$contact2 <- ifelse(data_Q3$contact == "cellular", 1, 0)

data_Q3$month1 <- ifelse(data_Q3$month == "jan", 1, 0)
data_Q3$month2 <- ifelse(data_Q3$month == "feb", 2, 0)
data_Q3$month3 <- ifelse(data_Q3$month == "mar", 3, 0)
data_Q3$month4 <- ifelse(data_Q3$month == "apr", 4, 0)
data_Q3$month5 <- ifelse(data_Q3$month == "may", 5, 0)
data_Q3$month6 <- ifelse(data_Q3$month == "jun", 6, 0)
data_Q3$month7 <- ifelse(data_Q3$month == "jul", 7, 0)
data_Q3$month8 <- ifelse(data_Q3$month == "aug", 8, 0)
data_Q3$month9 <- ifelse(data_Q3$month == "sep", 9, 0)
data_Q3$month10 <- ifelse(data_Q3$month == "oct", 10, 0)
data_Q3$month11 <- ifelse(data_Q3$month == "nov", 11, 0)

data_Q3$poutcome1 <- ifelse(data_Q3$poutcome == "unknown", 1, 0)
data_Q3$poutcome2 <- ifelse(data_Q3$poutcome == "other", 1, 0)
data_Q3$poutcome3 <- ifelse(data_Q3$poutcome == "failure", 1, 0)

data_Q3$y10 <- ifelse(data_Q3$y == "yes", 1, 0)

model_logreg <- glm(y10 ~ age + 
                      job1 + job2 + job3 + job4 + job5 + 
                      job6 + job7 + job8 + job9 + job10 + job11 + 
                      marital1 + marital2 + 
                      edu1 + edu2 + edu3 +
                      default10 + housing10 + loan10 + balance +
                      contact1 + contact2 + day + 
                      month1 + month2 + month3 + month4 + month5 + month6 +
                      month7 + month8 + month9 + month10 + month11 + duration + 
                      campaign + pdays + previous + 
                      poutcome1 + poutcome2 + poutcome3, data = data_Q3)

model2 <- step(model_logreg, direction = "backward")
```

## Question 3
Above model as an equation
```{r Question3.3}
model2$call[[2]]
```

## Question 4
Prediction accuracy
```{r Question3.4}
validation_Q3$job1 <- ifelse(validation_Q3$job == "admin.", 1, 0)
validation_Q3$job2 <- ifelse(validation_Q3$job == "unknown", 1, 0)
validation_Q3$job3 <- ifelse(validation_Q3$job == "unemployed", 1, 0)
validation_Q3$job4 <- ifelse(validation_Q3$job == "management", 1, 0)
validation_Q3$job5 <- ifelse(validation_Q3$job == "housemaid", 1, 0)
validation_Q3$job6 <- ifelse(validation_Q3$job == "entrepreneur", 1, 0)
validation_Q3$job7 <- ifelse(validation_Q3$job == "student", 1, 0)
validation_Q3$job8 <- ifelse(validation_Q3$job == "blue-collar", 1, 0)
validation_Q3$job9 <- ifelse(validation_Q3$job == "self-employed", 1, 0)
validation_Q3$job10 <- ifelse(validation_Q3$job == "retired", 1, 0)
validation_Q3$job11 <- ifelse(validation_Q3$job == "technician", 1, 0)

validation_Q3$marital1 <- ifelse(validation_Q3$marital == "married", 1, 0)
validation_Q3$marital2 <- ifelse(validation_Q3$marital == "divorced", 1, 0)

validation_Q3$edu1 <- ifelse(validation_Q3$education == "secondary", 1, 0)
validation_Q3$edu2 <- ifelse(validation_Q3$education == "primary", 1, 0)
validation_Q3$edu3 <- ifelse(validation_Q3$education == "tertiary", 1, 0)

validation_Q3$default10 <- ifelse(validation_Q3$default == "yes", 1, 0)

validation_Q3$housing10 <- ifelse(validation_Q3$housing == "yes", 1, 0)

validation_Q3$loan10 <- ifelse(validation_Q3$loan == "yes", 1, 0)

validation_Q3$contact1 <- ifelse(validation_Q3$contact == "telephone", 1, 0)
validation_Q3$contact2 <- ifelse(validation_Q3$contact == "cellular", 1, 0)

validation_Q3$month1 <- ifelse(validation_Q3$month == "jan", 1, 0)
validation_Q3$month2 <- ifelse(validation_Q3$month == "feb", 2, 0)
validation_Q3$month3 <- ifelse(validation_Q3$month == "mar", 3, 0)
validation_Q3$month4 <- ifelse(validation_Q3$month == "apr", 4, 0)
validation_Q3$month5 <- ifelse(validation_Q3$month == "may", 5, 0)
validation_Q3$month6 <- ifelse(validation_Q3$month == "jun", 6, 0)
validation_Q3$month7 <- ifelse(validation_Q3$month == "jul", 7, 0)
validation_Q3$month8 <- ifelse(validation_Q3$month == "aug", 8, 0)
validation_Q3$month9 <- ifelse(validation_Q3$month == "sep", 9, 0)
validation_Q3$month10 <- ifelse(validation_Q3$month == "oct", 10, 0)
validation_Q3$month11 <- ifelse(validation_Q3$month == "nov", 11, 0)

validation_Q3$poutcome1 <- ifelse(validation_Q3$poutcome == "unknown", 1, 0)
validation_Q3$poutcome2 <- ifelse(validation_Q3$poutcome == "other", 1, 0)
validation_Q3$poutcome3 <- ifelse(validation_Q3$poutcome == "failure", 1, 0)

validation_Q3$y10 <- ifelse(validation_Q3$y == "yes", 1, 0)
final_logreg <- glm(model2$call[[2]], data = data_Q3)
pred <- predict(final_logreg, validation_Q3)
pred <- ifelse(pred < 0.5, 0, 1)

acc <- sum(pred == validation_Q3$y10) / length(pred) * 100
paste0("Prediction accuracy = ", acc, "%")

```

