---
title: "DA5030.Proj.Parpattedar"
author: "Shruti Parpattedar"
date: "April 21, 2019"
output: slidy_presentation
---

# Data Acquisition
X1	Relative Compactness 
X2	Surface Area 
X3	Wall Area 
X4	Roof Area 
X5	Overall Height 
X6	Orientation 
X7	Glazing Area 
X8	Glazing Area Distribution 
y1	Heating Load 
y2	Cooling Load
```{r}
library(readxl)
proj_data <- data.frame(read_excel("ENB2012_data.xlsx"))
colnames(proj_data) <- c("Rel_Com", "Surf_Area", "Wall_Area", 
                         "Roof_Area", "Ov_Hght", "Orient", 
                         "Glaz_Area", "Glaz_A_Dist", "Heating", "Cooling")

str(proj_data)
```

# Data Exploration

```{r}
library(psych)
hist((proj_data$Rel_Com))
pairs.panels(proj_data)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

