---
title: "DA5030.Proj.Parpattedar"
author: "Shruti Parpattedar"
date: "April 21, 2019"
output: pdf_document
---

# Data Acquisition
X1	Relative Compactness 
X2	Surface Area - m2
X3	Wall Area - m2
X4	Roof Area - m2
X5	Overall Height - m 
X6	Orientation - 2:North, 3:East, 4:South, 5:West
X7	Glazing Area - 0%, 10%, 25%, 40% (of floor area)
X8	Glazing Area Distribution - 1:Uniform, 2:North, 3:East, 4:South, 5:West
y1	Heating Load - kWh/m²
y2	Cooling Load -  kWh/m²

Loading data from an excel sheet. Considering all features as continuous.
```{r}
library(readxl)
proj_data <- data.frame(read_excel("ENB2012_data.xlsx"))
colnames(proj_data) <- c("Rel_Com", "Surf_Area", "Wall_Area", 
                         "Roof_Area", "Ov_Hght", "Orient", 
                         "Glaz_Area", "Glaz_A_Dist", "Heating", "Cooling")

str(proj_data)
summary(proj_data)
```

# Data Exploration
## Exploratory data plots
Displaying histogram for various features.


Observations from ggplot - 
- Roof and surface area range is high when the overall height is low,
- Roof and surface area range is low when the overall height is high
- There are no observations for high overall height and high surface area and also for low overall height and low surface area.

## Outlier Detection
Considering 3 std dev from mean to be an outlier. Using this, no outliers were found.
```{r}
library(psych)
library(ggplot2)

# Displaying histograms for all attributes of the dataset
lapply(proj_data[,c(1:5, 9, 10)], hist)
hist(log(proj_data[,9]))

# Displaying pairwise scatterplots and correlation, and histograms
pairs.panels(proj_data)

# Observing relation between Roof Area, Surface Area and Glazing Area and how the Load is distributed using scatter plot
ggplot(proj_data, aes(x = Cooling, y = Heating), alpha = 0.3)+
  geom_point(aes(colour = Roof_Area ))+
  facet_grid(Ov_Hght + Glaz_Area ~ Surf_Area,  space = "free") +
  ggtitle("Load distribuiton of energy by Roof Area and Surface Area \n by Glazing Area and Overall Height")

# Function to detect outliers
outliers <- function(x) 
{
  for(i in 1:ncol(x))
  {
    sd_i <- sd(x[,i])
    mean_i <- mean(x[,i])
    
    out = x[x[,i] > 3*sd_i+mean_i | x[,i] < mean_i-3*sd_i, ]
    if(nrow(out) > 0)
    {
      print(colnames(x)[i])
      paste("The outliers are -", out)
    }else
    {
      print(paste("No outliers for",colnames(x)[i]))
    }
  }
}

# Detecting outliers in the project dataset
outliers(proj_data)
```

# Data Cleaning & Shaping
No NAs were found so adding NAs at random and then imputing them.

Normalizing the feature variables using min-max normalization.

Implementing principal component analysis.
```{r}
# Detecting NAs
proj_data[is.na(proj_data),]

# Adding NAs for data imputation since none of them already exist
# Adding 10 NAs in random positions
data_w_NAs = proj_data
for (i in 1:10) {
  row = sample(1:768, 1)
  col = sample(1:8, 1)
  data_w_NAs[row, col] = NA
}
#t <- aggregate(data = proj_data, Rel_Com ~ Surf_Area, mean, na.rm = TRUE)

getMode <- function(x) {
  uniq <- unique(x)
  uniq[which.max(tabulate(match(x, uniq)))]
}

for(i in 1:nrow(data_w_NAs))
{
  for(j in 1:10)
  {
    if(j == 6 | j == 7 | j == 8)
    {
      if(is.na(data_w_NAs[i,j]))
      {
        data_w_NAs[i, j] = getMode(data_w_NAs[,j])
      }
    }
    else
    {
      if(is.na(data_w_NAs[i,j]))
      {
        paste(data_w_NAs[i,j])
        data_w_NAs[i, j] = mean(data_w_NAs[,j], na.rm = TRUE)
      }
    }
  }
}

# Normalization function using min-max noramlization
normalize <- function(x) 
{
  return ((x - min(x)) / (max(x) - min(x))) 
}

# Normalizing the feature variables with continuous values
cont_v <- c(1:5)
data_norm <- cbind(normalize(proj_data[,cont_v]), proj_data[, c(6:10)])
data_norm_w_NA <- cbind(normalize(data_w_NAs[,cont_v]), data_w_NAs[, c(6:10)])

# PCA on the trained, scaled dataset
A1 = prcomp(data_norm[,1:5])

# Summary of the results
summary(A1)
plot(A1, type="l", main = "Principal Component Analysis")
```

# Model Construction & Evaluation
Cannot implement the hold-out method since unique combinations of data are observed in the dataset.
```{r}
# Creating training and validation datasets
# sample <- sample.int(n = nrow(data_norm), size = 0.7*nrow(data_norm), replace = FALSE)
# train_data <- data_norm[sample,]
# validn_data <- data_norm[-sample,]
```

## Building kNN regression model
Building a k Nearest Neighbors regression model to predict response values of the given data.
```{r}
# Regression version of kNN
kNN.reg <- function(new_data, target_data, train_data, k)
{
  n <- nrow(train_data)
  d <- rep(0,n)
  for (i in 1:n) 
  {
    d[i] <- sqrt(sum((train_data[i,1:8] - new_data[1:8])^2))
  }
  o <- order(d)
  m <- mean(target_data[o[1:k]]) 
  return(m)
}

for(i in 1: nrow(data_norm))
{
  data_norm$Heating_reg_kNN[i] <- kNN.reg(data_norm[i,], data_norm[,9], data_norm[,1:8], 10)
  data_norm$Cooling_reg_kNN[i] <- kNN.reg(data_norm[i,], data_norm[,10], data_norm[,1:8], 10)
  data_norm_w_NA$Heating_reg_kNN[i] <- kNN.reg(data_norm_w_NA[i,], data_norm_w_NA[,9], data_norm_w_NA[,1:8], 10)
  data_norm_w_NA$Cooling_reg_kNN[i] <- kNN.reg(data_norm_w_NA[i,], data_norm_w_NA[,10], data_norm_w_NA[,1:8], 10)
}

# data_norm_w_NA$Heating_reg_kNN <- reg_knn_heating_NA 
# data_norm_w_NA$Cooling_reg_kNN <- reg_knn_cooling_NA

new_data <- c(0.69, 735.0, 294.0, 220.50, 3.5, 4, 0.25, 4)
for(i in 1:5)
{
  new_data[i] <- (new_data[i] - min(proj_data[,i])) / (max(proj_data[,i]) -
                                                               min(proj_data[,i]))
}

# Predicting response values of new data
kNN.reg(new_data, data_norm$Heating, data_norm[,1:8], 10)
kNN.reg(new_data, data_norm$Cooling, data_norm[,1:8], 10)

```

# Multiple Regression
Implementation of multiple regression for the two response variables.
```{r}
# Model for heating 
model_heating <- lm(Heating ~ Rel_Com + Surf_Area + Wall_Area + Roof_Area + Ov_Hght +
                     Orient + Glaz_A_Dist + Glaz_Area, data = data_norm)

model_heating <- step(model_heating, direction = "backward")
summary(model_heating)
data_norm$Heating_lm <- predict(model_heating, data_norm)
data_norm_w_NA$Heating_lm <- predict(model_heating, data_norm_w_NA)

# Model for cooling
model_cooling <- lm(Cooling ~ Rel_Com + Surf_Area + Wall_Area + Roof_Area + Ov_Hght +
                     Orient + Glaz_A_Dist + Glaz_Area, data = data_norm)

model_cooling <- step(model_cooling, direction = "backward")
summary(model_cooling)
data_norm$Cooling_lm <- predict(model_cooling, data_norm)
data_norm_w_NA$Cooling_lm <- predict(model_cooling, data_norm_w_NA)

```

# Evaluation with k-fold cross-validation
Implementing k-fold cross-validation
```{r}
library(e1071)
# Heating
# Full Data set used for cross validation
knn.cross <- tune.knn(x = data_norm[,1:8], y = factor(data_norm[,9]), 
                      k = 2:20,tunecontrol=tune.control(sampling = "cross"), 
                      cross=10)

# Summarize the resampling results set
summary(knn.cross)
plot(knn.cross)

# Cooling
# Full Data set used for cross validation
knn.cross <- tune.knn(x = data_norm[,1:8], y = factor(proj_data[,10]), 
                      k = 2:20,tunecontrol=tune.control(sampling = "cross"), 
                      cross=10)

# Summarize the resampling results set
summary(knn.cross)
plot(knn.cross)
```

# Tuning the models
Tuning the kNN model with bootstrap sampling
```{r}
library(e1071)
# Heating
# Resampling using bootstrap
knn.boot <- tune.knn(x = proj_data[,1:8], y = factor(round(proj_data[,9])), 
                     k = 2:20,tunecontrol=tune.control(sampling = "boot"), 
                     boot=10)

# Summarize the resampling results set
summary(knn.boot)
plot(knn.boot)

# Cooling
# Resampling using bootstrap
knn.boot <- tune.knn(x = proj_data[,1:8], y = factor(round(proj_data[,10])), 
                     k = 2:20,tunecontrol=tune.control(sampling = "boot"), 
                     boot=10)

# Summarize the resampling results set
summary(knn.boot)
plot(knn.boot)
```

# Comparison using RMSE
Comparing the various models using RMSE.
```{r}
calc_RMSE <- function(orig, pred)
{
  sqrt(mean((orig-pred)^2))
}

RMSE_knn_heating <- calc_RMSE(data_norm$Heating, data_norm$Heating_reg_kNN)
RMSE_knn_cooling <- calc_RMSE(data_norm$Cooling, data_norm$Cooling_reg_kNN)
RMSE_lm_heating <- calc_RMSE(data_norm$Heating, data_norm$Heating_reg_kNN)
RMSE_lm_cooling <- calc_RMSE(data_norm$Cooling, data_norm$Cooling_reg_kNN)

crit = c("RMSE_knn_heating", "RMSE_knn_cooling", "RMSE_lm_heating", "RMSE_lm_cooling")
rmses = c(RMSE_knn_heating, RMSE_knn_cooling, RMSE_lm_heating, RMSE_lm_cooling)
Summary <- data.frame(Criteria = crit, RMSE = rmses)
Summary

RMSE_knn_heating_NA <- calc_RMSE(data_norm_w_NA$Heating, data_norm_w_NA$Heating_reg_kNN)
RMSE_knn_cooling_NA <- calc_RMSE(data_norm_w_NA$Cooling, data_norm_w_NA$Cooling_reg_kNN)
RMSE_lm_heating_NA <- calc_RMSE(data_norm_w_NA$Heating, data_norm_w_NA$Heating_reg_kNN)
RMSE_lm_cooling_NA <- calc_RMSE(data_norm_w_NA$Cooling, data_norm_w_NA$Cooling_reg_kNN)

crit = c("RMSE_knn_heating_NA", "RMSE_knn_cooling_NA", "RMSE_lm_heating_NA", "RMSE_lm_cooling_NA")
rmses = c(RMSE_knn_heating_NA, RMSE_knn_cooling_NA, RMSE_lm_heating_NA, RMSE_lm_cooling_NA)
Summary_NA <- data.frame(Criteria = crit, RMSE = rmses)
Summary_NA

```

