---
title: "DA5030.A6.Parpattedar"
author: "Shruti Parpattedar"
date: "March 11, 2019"
output: pdf_document
---
# Problem 1

## Question 1
Scatterplots and correlations between Age, absences, G1, G2 and G3 using pairs.panel().
```{r Question1.1}
library(psych)
data <- read.csv("student-mat.csv", sep = ";")
str(data)
pairs.panels(data[,c(3,30:33)])
```

## Question 2
Multiple regression model for the final grade, G3 using G2, G1, F0, F1, absences, schoolsup, health1, health2, internet.

The adjusted R-squared value suggests that the model covers approximately 83% of the variance in the data.
```{r Question1.2}
data$F0 <- ifelse(data$failures == 0, 1, 0) 
data$F1 <- ifelse(data$failures == 1, 1, 0) 
data$F2 <- ifelse(data$failures == 2, 1, 0) 
data$schoolsup <- ifelse(data$schoolsup == 'yes', 1, 0)
data$health1 <- ifelse(data$health == 1,1,0)
data$health2 <- ifelse(data$health == 2,1,0)
data$health3 <- ifelse(data$health == 3,1,0)
data$health4 <- ifelse(data$health == 4,1,0)
data$internet <- ifelse(data$internet == 'yes', 1,0)
m <- lm(G3 ~ G2+G1+F0+F1+absences+schoolsup+health1+health2+internet, data = data)
summary(m)
```

## Question 3
Using stepwise backward elimination to remove all non-significant variables from the model used in the previous question. 

The final model for the final grade, G3 uses G2, G1, F1, absences, schoolsup, health2 variables.

AIC used as a backward elimination measure.
```{r Question1.3}
m2 <- step(lm(G3 ~ G2+G1+F0+F1+absences+schoolsup+health1+health2+internet, data = data), 
           direction = "backward")

# Final model
m2$call[[2]]

# Equation to be used for prediction
# g3 <- m2$coefficients[[1]] + g2*m2$coefficients[[2]] + g1*m2$coefficients[[3]] 
# + f1*m2$coefficients[[4]] + abs*m2$coefficients[[5]]+ schsup*m2$coefficients[[6]]
# + h2*m2$coefficients[[7]]
```

## Question 4
Predicting the final grade for a new student with the following data and then calculating the 95% Confidence Interval for the prediction.
```{r Question1.4}
g2 <- 13
g1 <- 14
f1 <- 0
abs <- 3
schsup <- 0
h2 <- 0
new <- c(g2,g1,f1,abs,schsup,h2)
g3 <- m2$coefficients[[1]] + g2*m2$coefficients[[2]] + g1*m2$coefficients[[3]] + 
  f1*m2$coefficients[[4]] + abs*m2$coefficients[[5]] + schsup*m2$coefficients[[6]] +
  h2*m2$coefficients[[7]]

# Prediction for new student
g3

# Confindence interval for prediction
CI <- c(g3 - 1.96*1.877, g3 + 1.96*1.877)
CI
```

## Question 5
Calculating the RMSE for the multiple regression model
```{r Question1.5}
finalmodel <- lm(m2$call[[2]], data)
pred <- predict(finalmodel, data)
rmse1_lm <- sqrt(mean((data$G3 - pred)^2))
rmse2_lm <- sqrt(mean((finalmodel$residuals)^2))
paste("RMSE for Multiple Regression -", rmse2_lm)
```

# Problem 2

## Question 1
Adding a Pass/Fail column based on the G3 column.

Next, converting the new column into dummy codes.
```{r Question2.1}
data$PF <- ifelse(data$G3<10, 'F', 'P')
data$PF10 <- ifelse(data$PF == 'P', 1, 0)
```

## Question 2
Generating a logistic regression model for the Pass/fail dummy codes column using age, Medu, goout1, goout2, goout3, goout4, G1, G2, G3, F0, F1, absences, schoolsup.

Next, using stepwise backward elimination to remove all non-significant variables from the model.
```{r Question2.2}
data$goout1 <- ifelse(data$goout == 1, 1, 0)
data$goout2 <- ifelse(data$goout == 2, 1, 0)
data$goout3 <- ifelse(data$goout == 3, 1, 0)
data$goout4 <- ifelse(data$goout == 4, 1, 0)

model_logreg <- glm(PF10 ~ age + Medu + goout1 + goout2 + goout3 + goout4 + 
                      G1 + G2 + G3 + F0 +F1 + absences + schoolsup, data = data)
model_logreg2 <- step(model_logreg, direction = "backward")
```

## Question 3
Displaying the model with only the significant variables
```{r Question2.3}
model_logreg2$call[[2]]
```

## Question 4
Using the model to make predictions using the entire dataset. 

Since logistic regression models result in possibilities ie. continuous values, using a threshold of 0.5 to convert them in to binary values.

Calculating the RMSE for the logistic regression model.
```{r Question2.4}
final_logreg <- glm(model_logreg2$call[[2]], data = data)
pred <- predict(final_logreg, data)
pred <- ifelse(pred < 0.5, 0, 1)
rmse1_glm <- sqrt(mean((data$PF10 - pred)^2))
rmse2_glm <- sqrt(mean(round(final_logreg$residuals)^2))
paste("RMSE for Logistic Regression -", rmse2_glm)
```

# Problem 3

## Question 1

### Step 1 - Collecting Data
```{r Question3.1.1}
wine <- read.csv("whitewines.csv")
```

### Step 2 - Exploring and preparing the data 
Creating the train and test datasets.
```{r Question3.1.2}
str(wine)
hist(wine$quality) 
wine_train <- wine[1:3750, ] 
wine_test <- wine[3751:4898, ]
```

### Step 3 - Training a model on the data
Using the rpart function with the training data.

Next, using rpart.plot to viualize the model.
```{r Question3.1.3}
library(rpart)
library(rpart.plot) 
m.rpart <- rpart(quality ~ ., data = wine_train)
m.rpart

# Visualizing decision trees
rpart.plot(m.rpart, digits = 3)

rpart.plot(m.rpart, digits = 4, fallen.leaves = TRUE, type = 3, extra = 101)
```

### Step 4 - Evaluating model performance 
Using the model to make predictions.

Next, finding the Mean Absolute Error for the model and its predictions
```{r Question3.1.4}
p.rpart <- predict(m.rpart, wine_test)
summary(p.rpart)
summary(wine_test$quality)
cor(p.rpart, wine_test$quality)

# Measuring performance with the mean absolute error 
MAE <- function(actual, predicted) 
{
  mean(abs(actual - predicted)) 
} 

MAE(p.rpart, wine_test$quality)

mean(wine_train$quality)
MAE(5.89, wine_test$quality)
```

### Step 5 - Improving model performance
Using the M5P function from the Rweka package.

Next, finding the MAE for this model.
```{r Question3.1.5}
library(RWeka)
m.m5p <- M5P(quality ~ ., data = wine_train)
m.m5p
summary(m.m5p)
p.m5p <- predict(m.m5p, wine_test)
summary(p.m5p)
cor(p.m5p, wine_test$quality)
MAE(wine_test$quality, p.m5p)
```

## Question 2
RMSEs for the two models used above.
```{r Question3.2}
rmse_rpart <- sqrt(mean((wine_test$quality - p.rpart)^2))
paste("RMSE for Rpart model -", rmse_rpart)
rmse_m5p <- sqrt(mean((wine_test$quality - p.m5p)^2))
paste("RMSE for M5P model - ", rmse_m5p)
```
